# Calculate Mussel Metrics

## Prepare Mussel Sample Metrics
The purpose of this section is to calculate various metrics of habitat quality for each sampled site. Each of these metrics will be used as presence data as input to create a set of Maxent habitat suitability models. The metrics calculated in this section were taken from the Upper Mississippi River Mussel Community Assessment Tool (MCAT) [@dunn2016].

## MCAT Metrics
The following MCAT metrics were selected for inclusion to the model:

* `% listed species`: percent of listed threatened or endangered species (federal or bordering states) is a measure of sensitive species
* `% tolerant`: percent of tolerant species (Amblema plicata, Quadrula quadrula, and Obliquaria reflexa) is a measure of a highly disturbed mussel assemblage; i.e., dominated by species tolerant of unstable substrates, silt accumulation, low current velocities, and fluctuating flow conditions.
* `% tribe Lampsilini`: percent of assemblage that falls within tribe Lampsilini is a measure of species composition, life history, and behavioral characteristics
* `% juveniles`: percent of mussels <= 5 years-old is a measure of recruitment into an assemblage over the past five years
* `% >= 15 yrs`: percent of mussels >= 15 years-old is a measure of age distribution in an assemblage
* `Q75 abundance`: a measure of abundance at the 75th quartile calculated as the density (no./m^2)
* `species evenness`: species evenness represents the dominance of an assemblage by a few species using Pielou's evenness index (range 0 to 1)
* `tribe evenness`: tribe evenness represents the dominance of a particular taxonomic group using Pielou's evenness index (range 0 to 1)
* `ES_100`: the expected number of species with a sample size of 100 estimated by rarefaction is a measure of a healthy mussel assemblage

## Create the Individuals table
The purpose of this step is to assemble the "individual" data for each of the surveys within the study area that will be used in the study. 

* Working geodatabase: `//mvrdfs/EGIS/Work/EMP/HREP_Projects/SteamboatSlough/Mussels/SteamboatMusselModel/Data/SB_Mussels.gdb/SB_ILStatePlaneW`
* Create a feature class respresenting all individual mussel samples in the study area, using a definition query to include only live individuals and name it `Cordova_all_individuals`. Note: the Q75 abundance metric will be based on live mussel counts for each quadrat.
* Add the following fields (Type=long integer) to the attribute table: `Listed`, `Tolerant`, `Lampsilini`, `Juveniles`, `Over_15yrs`, `tribe` 

## Create Percent Listed
The purpose of this step is to calculate the percent listed MCAT metric for each sampled site. 

* Dissolve this feature class by `SAMPLE_ID` and `ENAME` (Scientific name).  Add 'SUM_NumberLive' and 'SUM_Listed' fields in the Statistics section of the tool. Set Output Feature Class to the working .gdb and name it `PercentListed`.
* Open the `PercentListed` attribute table. Add a field titled 'Perc_listed' Type=double and enter the expression `([SUM_Listed] / [SUM_NumberLive]) * 100` into the field calculator to calculate % of individual listed species per quadrat sample.
* Add 3 fields titled 'species' Type=string and Length=20; 'longitude' Type=double; and 'latitude' Type=double; in that order. 
* Use field calculator to populate the field with species="perc_listed". Use Calculate Geometry to populate the 'longitude' and 'latitude' fields, ensuring that the coordinate system is set to 'NAD 1983 2011 US Feet' and 'StatePlane Illinois West FIPS 1202' and the units are set to US feet.
* Use the Symbology tab to select a suitability threshold based on quartiles (e.g., only selecting "fair" or "good" locations with >=50% listed species). In the Symbology tab, Show: Quantities; field value 'perc_listed'; classes=5; and method=quantile. Export selected to .csv for Maxent model run.

## Create Percent Tolerant
The purpose of this step is to calculate the percent tolerant MCAT metric for each sampled site. 

## Create Percent Tribe Lampsilini
The purpose of this step is to calculate the percent tribe Lampsilini MCAT metric for each sampled site.

## Create Percent Juveniles
The purpose of this step is to calculate the percent juveniles MCAT metric for each sampled site.

## Create Percent Over 15 Years
The purpose of this step is to calculate the percent over 15 years old MCAT metric for each sampled site.

## Create Q75 Abundance
The purpose of this step is to calculate the Q75 abundance MCAT metric for each sampled site.

* For Q75 abundance metric, dissolve `Cordova_all_individuals` by `SAMPLE_ID` and `ENAME` (Scientific name).  Add 'SUM_NumberLive' field in the Statistics section of the tool. Set Output Feature Class to the working .gdb and name it `Q75_abundance`.
* Use the Symbology tab to select a suitability threshold based on quartiles (e.g., only selecting the upper quartiles with >=50% live species per quadrat). Export selection to .csv for Maxent model run.

## Create Species Evenness
The purpose of this step is to calculate the Pielou's evenness index (range 0-1), estimated at the species level for each sampled site. 

* Create a table of SampleID by species by abundance (number live). Use the `Dissolve` tool:
  * Input Features: `Cordova_all_individuals`
  * Output feature class: `Cordova_sampleid_species`
  * Dissolve Fields: `SampleID`, `EName`
  * Statistics Fields: `NumberLive`, SUM
  * Create Multipart: unchecked

Ensure that the `Cordova_sampleid_species` feature class is of type `Point Features` and not `Multipoint Features`. If multipoint, then use the `Multipart to Singlepart` tool to convert from multipoint to point feature type. 

```{r message=FALSE, warning=FALSE, include=FALSE}
# Load libraries
library(sp)
library(labdsv)
library(vegan)
library(ggplot2)
library(dplyr)
library(arcgisbinding)

# Check for ArcGIS license
arc.check_product()

# Define some ArcGIS helper functions

# Converts ArcGIS feature class to an sp object
arc2sp <- function(fc_path) {
  # Open a connection to the specified ArcGIS feature class
  arcobj <- arcgisbinding::arc.open(fc_path)
  # Select the ArcGIS data
  arc <- arcgisbinding::arc.select(arcobj)
  # Convert the ArcGIS format to the sp format
  sp <- arcgisbinding::arc.data2sp(arc)
}

# Converts an sp object to an ArcGIS feature class
sp2arc <- function(sp_object, fc_path) {
  # Convert the sp object to an ArcGIS data object
  arcobj <- arcgisbinding::arc.sp2data(sp_object)
  # Write the ArcGIS object to a geodatabase feature class
  arcgisbinding::arc.write(data = arcobj, path = fc_path)
}
```

```{r}
# Import ArcGIS feature class into R
cordova_sampleid_species <- arc2sp("//mvrdfs/EGIS/Work/EMP/HREP_Projects/SteamboatSlough/Mussels/SteamboatMusselModel/Data/SB_Mussels.gdb/SB_ILStatePlaneW/Cordova_sampleid_species")

# Convert sp object to a data frame
sample_species <- cordova_sampleid_species@data

# Remove OBJECTID field
sample_species <- sample_species[,-1]

# Convert to vegan community data matrix-like format using labdsv::matrify
sample_species_matrix <- labdsv::matrify(sample_species)
```

```{r}
# Calculate the Shannon-Weaver diversity index
shannon_diversity <- vegan::diversity(sample_species_matrix, index = "shannon")

# Calculate the number species per site (see ?vegan::diversity)
species_number <- vegan::specnumber(sample_species_matrix)

# Use the following equation to calculate Pielou's evenness
# See the vegan Diversity Vignette for details
pielou_evenness <- shannon_diversity/log(species_number)

# Set NaN pielou_evenness values to zero
pielou_evenness <- ifelse(is.nan(pielou_evenness), 0, pielou_evenness)
# Create a data frame of results
pielou <- data.frame(sampleid = names(pielou_evenness), 
                     shannon_diversity, 
                     species_number, 
                     pielou_evenness)
```

```{r}
# Merge the Pielou scores back onto the sp object
sample_species_pielou <- sp::merge(x = cordova_sampleid_species,
                                   y = pielou,
                                   by.x = "SampleID", by.y = "sampleid")
```

```{r eval=FALSE}
# Export to geodatabase
sp2arc(sample_species_pielou, fc_path = "//mvrdfs/EGIS/Work/EMP/HREP_Projects/SteamboatSlough/Mussels/SteamboatMusselModel/Data/SB_Mussels.gdb/SB_ILStatePlaneW/sample_species_pielou")
```

* Create a table of SampleID by species Pielou evenness score. Use the `Dissolve` tool:
  * Input Features: `sample_species_pielou`
  * Output feature class: `species_pielou`
  * Dissolve Fields: `SampleID`
  * Statistics Fields: `shannon_diversity`, MEAN
                       `species_number`, MEAN
                       `pielou_evenness`, MEAN
  * Create Multipart: unchecked

* Add "SWD" fields.

## Create Tribe Eveness
The purpose of this step is to calculate the Pielou's evenness index (range 0-1), estimated at the tribe level.  (will probably need to add a new column for tribe; dissolve on this with sample ID)

* In the feature class `Cordova_all_individuals`, create a new text variable named `tribe`. 
* Use the `Field Calculator` tool on the `tribe` field to calculate its value using the following Python expression: `!Ename!.split()[0]`
* Create a table of SampleID by tribe by abundance (number live). Use the `Dissolve` tool:
  * Input Features: `Cordova_all_individuals`
  * Output feature class: `Cordova_sampleid_tribe`
  * Dissolve Fields: `SampleID`, `tribe`
  * Statistics Fields: `NumberLive`, SUM
  * Create Multipart: unchecked


```{r}
# Import ArcGIS feature class into R
cordova_sampleid_tribe <- arc2sp("//mvrdfs/EGIS/Work/EMP/HREP_Projects/SteamboatSlough/Mussels/SteamboatMusselModel/Data/SB_Mussels.gdb/SB_ILStatePlaneW/Cordova_sampleid_tribe")

# Convert sp object to a data frame
sample_tribe <- cordova_sampleid_tribe@data

# Remove OBJECTID field
sample_tribe <- sample_tribe[,-1]

# Convert to vegan community data matrix-like format using labdsv::matrify
sample_tribe_matrix <- labdsv::matrify(sample_tribe)
```


```{r}
# Calculate the Shannon-Weaver diversity index
tribe_shannon_diversity <- vegan::diversity(sample_tribe_matrix, index = "shannon")

# Calculate the number species per site (see ?vegan::diversity)
tribe_number <- vegan::specnumber(sample_tribe_matrix)

# Use the following equation to calculate Pielou's evenness
# See the vegan Diversity Vignette for details
tribe_pielou_evenness <- tribe_shannon_diversity/log(tribe_number)

# Set NaN pielou_evenness values to zero
tribe_pielou_evenness <- ifelse(is.nan(tribe_pielou_evenness), 0, 
                                tribe_pielou_evenness)
# Create a data frame of results
tribe_pielou <- data.frame(sampleid = names(tribe_pielou_evenness), 
                           tribe_shannon_diversity, 
                           tribe_number, 
                           tribe_pielou_evenness)
```

```{r}
# Merge the Pielou scores back onto the sp object
sample_tribe_pielou <- sp::merge(x = cordova_sampleid_tribe,
                                 y = tribe_pielou,
                                 by.x = "SampleID", by.y = "sampleid")
```

```{r eval=FALSE}
# Export to geodatabase
sp2arc(sample_tribe_pielou, fc_path = "//mvrdfs/EGIS/Work/EMP/HREP_Projects/SteamboatSlough/Mussels/SteamboatMusselModel/Data/SB_Mussels.gdb/SB_ILStatePlaneW/sample_tribe_pielou")
```

* Create a table of SampleID by tribe Pielou evenness score. Use the `Dissolve` tool:
  * Input Features: `sample_tribe_pielou`
  * Output feature class: `tribe_pielou`
  * Dissolve Fields: `SampleID`
  * Statistics Fields: `tribe_shannon_diversity`, MEAN
                       `tribe_number`, MEAN
                       `tribe_pielou_evenness`, MEAN
  * Create Multipart: unchecked

* Add "SWD" fields.

## Create ES 100
The purpose of this step is to calculate the ES 100 MCAT metric for each sampled site.

ES_100-species richness estimated by rarefaction; caveat: sites need to be compared based on an equal sample size because # of species and # of individuals sampled are large correlated.  Not sure how best to approach this one; I defer to you!


## Determine MCAT Metric per site threshold
The purpose of this section is to display the range of calculated values for each of the MCAT metrics for each of the sites in the Steamboat Island study area. The MCAT report is clear about metric thresholds to use for evaluating quality mussel beds. However, this study is using the individual site unit of analysis. Therefore, decisions must be made on what is the appropriate metric threshold to use for the site level of analysis. The descriptive statistics in this section will help to inform that decsion. 

```{r message=FALSE, warning=FALSE, include=FALSE}
# Import MCAT Metrics from geodatabase
# Percent Listed
percent_listed_sp <- arc2sp("//mvrdfs/EGIS/Work/EMP/HREP_Projects/SteamboatSlough/Mussels/SteamboatMusselModel/Data/SB_Mussels.gdb/SB_ILStatePlaneW/PercentListed")

# Convert sp object to a data frame
percent_listed <- percent_listed_sp@data

# Percent Tolerant
percent_tolerant_sp <- arc2sp("//mvrdfs/EGIS/Work/EMP/HREP_Projects/SteamboatSlough/Mussels/SteamboatMusselModel/Data/SB_Mussels.gdb/SB_ILStatePlaneW/PercentTolerant")

# Convert sp object to a data frame
percent_tolerant <- percent_tolerant_sp@data
```

### Percent Listed
```{r}
# Plot percent listed
percent_listed_p <- ggplot(percent_listed, aes(Perc_listed)) + 
  geom_histogram(binwidth = 10)
print(percent_listed_p)

#sort(percent_listed$Perc_listed[percent_listed$Perc_listed > 0])
```

### Percent Tolerant
```{r}
# Plot percent listed
percent_tolerant_p <- ggplot(percent_tolerant, aes(Perc_tolerant)) + 
  geom_histogram(binwidth = 10)
print(percent_tolerant_p)

#sort(percent_tolerant$Perc_tolerant[percent_tolerant$Perc_tolerant < 50])
```

##  Choice of Site Level MCAT Metric Thresholds
The discriptive statistics presented in the previous section were used to select MCAT metrics thresolds. 


Table 1.


## Export to Maxent "SWD" format




